---
title: "hw_09"
author: "Haley Dylewski"
date: "10/26/2020"
output: github_document

---


Practice in Functions and Programming


________________________________________________________

Write functions that compute the variance and skewness of  numeric vector

________________________________________________________

This function calculates variance of a vector

Given the equation for variance is
Var(x) = (1/n-1)*sum((xi-X)^2, i=1 to n)
where X is the mean and n is the number of data points
```{r}
library(tidyverse)

variance <- function(x){
n <- length(x)
X_bar <- mean(x)
var_sum <- 0
i <- 1
for (i in seq_along(x)){
  var_sum <- var_sum + ((x[i]-X_bar)^2)
}
Var <- (1/(n-1))*var_sum
Var
}

example <-c(1,2,3,4,5,6,7)
p <- variance(example)
p
```

This function calculates the skew of a vector

Given the equation for skew is
skew = ((1/n-2)*sum((x1-X)^3, i=1 to n))/(var(x)^3/2)
```{r}
skew_calculator <- function(x){
n <- length(x)
X_bar <- mean(x)
X_bar
sk_sum <- 0
i <- 1

for (i in seq_along(x)) {
  sk_sum <- sk_sum + (x[i]- X_bar)^3
}

var <- variance(x)
skew <- ((1/n-2)*sk_sum)/(var^(3/2))

}
```

Sample data set and function calls

```{r}
sample <- c(1,2,3,4,5,6,7,8,9,10)
sample_var <- variance(sample)
sample_var
sample_skew <- skew_calculator(sample)
sample_skew

```

_______________________________________________________

Write both_na function that takes two vectors of the same length and returns the number of position in both that have an NA.
________________________________________________________

both_na()


```{r}
NA_counter <- function(x,y){
x_NA <-sum(is.na(x))
y_NA <-sum(is.na(y))
print(paste0("Total NA in vector 1 = ", x_NA))
print(paste0("Total NA in vector 2 = ", y_NA))
}

```
Sample data and function call


```{r}
N <- 10
x_num <- round(rnorm(N, 0, 5))
x_num[rbinom(N, 1, 0.2) == 1] <- NA 
x_num


M<-10
y_num <- round(rnorm(N, 0, 5))
y_num[rbinom(M, 1, 0.2) == 1] <- NA 
y_num

NA_counter(x_num,y_num)
```


________________________________________________________

Write a function to create a specific kind of plot, with useful labels that can take a data frame as an argument


function for plotting flight data. The data to plot is the only variable passed in. The function code can be easily changed for repurposing. 
________________________________________________________
```{r}
flight_plot <- function (in_data){
  x_axis <- in_data$month
  x_title <- in_data$number_of_flights
  x_lim <- c(1:12,1)
  treatement<- in_data$dest
  y_title <- "Month"
  y_axis <- "Number of flights"


  
ggplot(top_dest, aes(x = month, y = number_of_flights, color = treatement)) + 
  geom_line() + geom_point() + 
  xlim(1,12) + scale_x_discrete(name = y_title, limits = x_lim) + scale_y_continuous(name = y_axis, labels=waiver())
}


```
Sample data and function call
```{r}

dest_data <- nycflights13::flights %>%
  group_by(dest,month)%>%
  summarize(number_of_flights = n())
top_dest <- filter(dest_data, dest == "ORD" | dest == "ATL" | dest == "LAX")

flight_plot(top_dest)
```

_______________________________________________________

Write a function that takes one argument and does the following:
if the argument is a numeric vector, multiplies every number in the vector by 2
if the argument is a character vector, sorts the vector into alphabetical order
if the argument is anything else, issues an informative warning and returns the argument unchanged

_______________________________________________________


```{r}

num_vect <- c(1,2,3)
char_vect <- c('z','a','b','c')
bool_vect <- c(TRUE, FALSE, TRUE)

x <- bool_vect

vector_mod <- function(x){
class <- class(x)

if (class == "numeric"){
  x = x*2
  
}else if (class == "character"){
  x <- sort(x)
  
}else{
 warning("WARNING: Vector of invalid class. Must be of character or numerical class")
  x <- x
}

}

```
Sample data and function call

```{r}

print("Modified numerical: ")
num.vect <- c(1,2,3)
num.mod <- vector_mod(num_vect)
num.mod

print("Modified character:")
char.vect <- c('z','a','b','c')
char.mod <- vector_mod(char_vect)
char.mod

print("Modified logical: ")
bool.vect <- c(TRUE, FALSE, TRUE)
logi.mod <- vector_mod(bool_vect)
logi.mod
```
_______________________________________________________

Write a function that tries to calculate the mean of a vector using mean(). If that is successful, it returns the mean of the vector. If mean() throws an error, it issues an informative message and returns the input intact.

This function uses tryCatch to identify errors/warnings, if none are encountered it returns the completed calculation of the passed argument as X_bar
_______________________________________________________

```{r}

error_catcher <- function(code) {
  tryCatch(code,
    error = function(c) "error: Mean cannot be calculated",
    warning = function(c) "warning: Argument is not numeric or logical",
    finally = function(c) X_bar <- code
  )
}
```

Test condition and function call

```{r}
num_vect <- c(1,2,3)
char_vect <- c('z','a','b','c')
bool_vect <- c(TRUE, FALSE, TRUE)


error_catcher(mean(num_vect))
error_catcher(mean(char_vect))
error_catcher(mean(bool_vect))
```
Note that this function does not stop the averaging of logical since these are stored in memory as integers and so do not throw a warning or error. 

_______________________________________________________

Use system.time() (or, if you want to be more advanced about it, the bench or microbenchmark packages) to compare how long it takes to add two, 1-million-element random vectors and store the result in a third vector: a. Using the vectorized sum() operation 
b. Using a for loop with a pre-allocated vector to store the results 
c. Using a for loop without pre-allocating a vector to store the results

_______________________________________________________

Create vectors and call microbenchmark library

```{r}
library(microbenchmark)

v1 <- runif(n = 1000000, min = 1, max = 100)
v2 <-runif(n = 1000000, min = 1, max = 10)


```


Part a: using vectorized sum().
```{r}
vec_sum <- NA
microbenchmark(
  vec_sum[1] <- sum(v1),
  vec_sum[2] <- sum(v2))



```

Part b: Using a for loop with a pre-allocated vector to store the results 

```{r}
premade.vec <- vector(mode = "integer", length = 2)

microbenchmark(
for (i in 1:1000000){
  premade.vec[1] = premade.vec[1]+v1[i]
  premade.vec[2] = premade.vec[2]+v2[i]
}
)
premade.vec
```


Part 3: Using a for loop without pre-allocating a vector to store the result


```{r}
new.vec <- NA
microbenchmark(
for (i in 1:1000000){
  new.vec[1] = new.vec[1]+v1[i]
  new.vec[2] = new.vec[2]+v2[i]
}
)

```

Using vectorized sum is much faster than the for loop. Without preallocating memory, the computation is much much slower. 



