---
title: "hw_10"
author: "Haley Dylewski"
date: "10/26/2020"
output: github_document

---

This document includes fundamental exploration in Linear models and objects and methods.

_____________________________________________________

Linear Models

____________________________________________________

Part1: Using the diamonds dataset, create a linear model of price as a function of the weight of the diamond (carat).
```{r}
library(tidyverse)


carat_weight_mod <- lm(price ~ carat, data = diamonds)

cost_1carat <- carat_weight_mod$coefficients[2]*1 + carat_weight_mod$coefficients[1]

print(paste0("The cost of one carrot is estimated at $", round(cost_1carat, digits = 2)))

```


Part 2: Now create a linear model of price as a function of carat and color.
What difference does it make if you convert color to a character vector before you make the model? Note that this is probably what you want to do in general, unless you really know what you’re doing.

model without conversion to character first
```{r}
color_mod <- lm(price ~ carat + color, data = diamonds)

```

After conversion to character

```{r}
ch.color <- as.character(diamonds$color)
color_mod_chr <- lm(price ~ carat + ch.color, data = diamonds)

```
The linear regressions change significantly when not converted to a character. This is likely due to the color being stored as a factor and affecting the calculation



Part 3: Make some plots of the raw data, and of the model fits, to make an argument as to which model is more useful.
Bonus: Should the data have been pre-processed before making these models?


Price vs carrot model and data
```{r}
b_price <- carat_weight_mod$coefficients[1]
m_price <- carat_weight_mod$coefficients[1]

ggplot(diamonds, aes(carat, price)) +
  geom_abline(aes(intercept = b_price, slope = m_price  )) +
  geom_point()
 

```


________________________________________________________

Objects and methods

________________________________________________________

Part 1: Write a function that returns some kind of structured result, which you define as a new S3 class. Your class doesn’t need to be anything fancy, but it should be different from existing classes.
________________________________________________________


The below structure function will store data from a potentiostat(EIS reading) in my lab and add the technician, date, and test condition.

Starting with just one attribute:
```{r}
new_EIS_data <- function(df, condition){
  
  stopifnot(is.data.frame(df))
  stopifnot(is.character(condition))
  attr(df, "condition") <- condition
  structure(df, class = c("EIS_date", "data.frame"))

}
```
Test the constructor function with imported data
Import data
```{r}
EIS_raw_data <- read_csv('C:/Users/hdyle/Documents/Micr_590/reproducible_data_analysis/data/EIS.csv')
glimpse(EIS_raw_data)
```


```{r}
new_EIS_obj <- new_EIS_data(EIS_raw_data, condition = "Control")
glimpse(new_EIS_obj)
```
Looks pretty fine. Now adding in other attributes I ran into some  problems. Please find at the end of document.


________________________________________________________

Part 2: Write print and summarize methods for this class. These methods don’t need to be particularly useful, but they should be different from the default methods.

________________________________________________________

Summarize method:
```{r}


summary.EIS_data <- function(EIS_obj){
  df_sum <- summarize.data.frame(inst_obj)
  print(paste0("Condition: ", attr(EIS_obj, "condition"), sep = " "))
  df_sum
}

summary(new_EIS_obj)
```


Print method
```{r}
print.EIS_data <- function(EIS_obj){

  print(paste0("Condition: ", attr(EIS_obj, "condition"), sep = " "))
  print.data.frame(EIS_obj)
}

print.EIS_data(new_EIS_obj)

```

________________________________________________________

_______________________________________________________
Constructor with Multiple attributes *** HELP!!!***
I tried to take this assignment further by adding additional attributes, but encountered the below error. If you have any feedback on how to acheive this I would appreciate it. 
## {r}
new_EIS_data <- function(df, technician,condition){
  stopifnot(is.data.frame(df))
  stopifnot(is.character(technician))
  stopifnot(is.character(condition))
  
  attr(df,"technician", "condition") <- c(technician, condition)
  
  structure(df, class = c("EIS_date", "data.frame"))

#}


another_EIS_obj <- new_EIS_data(EIS_raw_data, technician = "Haley", condition = "Control")
glimpse(another_EIS_obj)
```

I was hoping to create multiple attributes but could not get it to work. I would love some feedback on this! 





