---
title: "hw_10"
author: "Haley Dylewski"
date: "10/26/2020"
output: github_document
---

_____________________________________________________

Linear Models

____________________________________________________

Part1: Using the diamonds dataset, create a linear model of price as a function of the weight of the diamond (carat).
```{r}
library(tidyverse)


carrot_weight_mod <- lm(price ~ carat, data = diamonds)

cost_1carrot <- carrot_weight_mod$coefficients[2]*1 + carrot_weight_mod$coefficients[1]
print(paste0("The cost of one carrot is estimated at $", round(cost_1carrot, digits = 2)))

```
Part 2: Now create a linear model of price as a function of carat and color.
What difference does it make if you convert color to a character vector before you make the model? Note that this is probably what you want to do in general, unless you really know what you’re doing.

model without converion to character first
```{r}
lm(price ~ color, data = diamonds)


```

After conversion to character

```{r}
ch.color <- as.character(diamonds$color)
lm(price ~ ch.color, data = diamonds)

```
The linear regressions change significantly when not converted to a character. This is likely due to the color being stored as a factor, changing the calculation

```{r}

```



