---
title: "stats_oneway_only"
author: "Haley Dylewski"
date: "11/29/2020"
output: html_document
---

---
title: "Project test code"
author: "Haley Dylewski"
date: "11/22/2020"
output: html_document
---


This is the Statistical analysis code that will be directly added to the app. It is missing th 2-way ANOVA which has some bugs

```{r}
library(tidyverse)
library(rstatix)
library(ggpubr)
library(viridis)
library(emmeans)
library(DescTools)
```


```{r}
sample_data <- read.csv("C:/Users/hdyle/Documents/Micr_590/reproducible_data_analysis/homework/final_project/sampledata_mod.csv", check.names = FALSE )
sample_data
```

For a one way/ pairwise test
```{r}
sample_data <- mutate_if(sample_data, is.character, as.factor)



stat_data <- pivot_longer(sample_data,
                          cols = !contains('Sample'),
                          names_to = "Time_points",
                          values_to = "Number")

```

____________________________________________________________________

Statistical Analysis for incorporation:
____________________________________________________________________




These will be replaced with user inputs
```{r}
groupings <- c("Sample", "Treatment")

input_options <- c("None","T-Test","One Way ANOVA")
input <- input_options[3]
post_hoc <- "Tukey"
```


```{r}
### Format data



    long_data <- pivot_longer(sample_data,
                          cols = !contains('Sample'),
                          names_to = "Time_points",
                          values_to = "Number")
  


### Convert all characters to factors. This is required for the stats to work
data_set <- mutate_if(long_data, is.character, as.factor)
str(data_set)



### Check for outliers and gt summary statistics. This includes mean and standard deviation



  outliers <-  data_set %>% 
    group_by(Time_points,Sample)%>%
    identify_outliers(Number)
  sum_stats <- data_set %>%
    group_by(Time_points,Sample) %>%
    get_summary_stats(Number, type = "mean_sd")
 sum_stats

#returns values for assessing pass or fail of these conditions
#shapirores_result
#shapirogn_result
#levene_result
#if TRUE, passed, if FALSE, failed





### Check normality assumptions
### 
### returns 2 values shapiro_result (normality) and 
###levene_result (homogeneity of variances) that will determine the 
### type fo statistical test you can perform
 
  
 ### Check normality of each group individually
 ### Use when few groups with many points
 ### Here the cutoff is 20 data points per group
 data_set_length <- length(data_set[[1]])
 if (data_set_length >= 20){
  group_norm <- data_set%>%
    group_by(Sample) %>%
    shapiro_test(Number) 
  gncheck <- group_norm$p
  for (i in seq_along(gncheck)){
    if (gncheck[i] >0.05){shapiro_result = TRUE
    }else {
      shapiro_result = FALSE
      break
      }
    } 
  }else {
  ### Check normality of all groups together by analyzing model
  ### residuals. 
  ### Use for many groups with few points each. Here it is <20


  model <- lm(Number ~ Sample, data = data_set)
  shapiro <- shapiro_test(residuals(model))  
  scheck <- shapiro$p.value
  for (i in seq_along(scheck)){
    if (scheck[i] >0.05){shapiro_result = TRUE
    }else {
      shapiro_result = FALSE
      break
      }
    }
  }
 ### homogeneity of variances
  ## if this is not met, run Welch one-way anova welch_anova_test()
  ## and must run Games-Howell post hoc or pairwise t-test
  ## games_howell_test(y~x)
  levene <- data_set%>% 
    levene_test(Number ~ Sample) 
  levcheck <- levene$p
  for (i in seq_along(levcheck)){
   if (levcheck[1] >0.05){lev_result = TRUE
   }else {lev_result = FALSE
          break
   }
  }
  


###Perform statistical analysis


### a function to call later that checks for significance after the
### test is run  
significance_check <- function(x){
  sig = FALSE  
  for (i in seq_along(x)){
       if(x[i] == '*'){
          sig = TRUE
        }
    }
  return(sig)
}

### The value for significance is initially set to FALSE to prevent
### any accidental carryover from previous calculations
significance = FALSE


 
## Do T-test or ANOVA

if(input == "None"){
  print("No statistical analysis preformed")
  
  } else if(input == "T-Test"){
    print("Pairwise student's T-test with p-value adjusted using Bonferroni method")
    stat.test <- data_set%>%
      group_by(Time_points)%>%
      pairwise_t_test(
      Number ~ Sample, 
       p.adjust.method = "bonferroni"
    )
    pwc <- stat.test
    
     
  } else if(input == "One Way ANOVA"){
    
    ## perform normal anova if levin is true
    if(levene_result == TRUE){
      stat.test <- data_set %>% 
        group_by(Time_points) %>%
        anova_test(Number ~ Sample)
      stat.test
     ## determine if significant
      stat_check <- stat.test[[7]]
      stat_check
      significance <- significance_check(stat_check)  
    
      ## post hoc given significant ANOVA
    if (significance == TRUE){
          ## Tukey comparisons
           pwc <- data_set %>% 
           group_by(Time_points) %>%
           tukey_hsd(Number ~ Sample)

      } else { print("No significance")} 
    
  ## if levene test was failed, use Welch anova
    }else if ( levene_result == FALSE){
      stat.test <- data_set %>%
        group_by(Time_points)%>%
        welch_anova_test(Number ~ Sample)
     ## determine if significant
      stat_check <- stat.test[[7]]
      significance <- significance_check(stat_check)  
      
      ## Games-Howell post hoc given significant ANOVA
      if (significance == TRUE){
       pwc <- data_set %>% 
         group_by(Time_points)%>%
         games_howell_test(Number ~ Sample)
      } else { print("No significance")} 
    }
          

} 



```

Return results

```{r}
print("Test")
input
print("shapiro")
shapiro_result
print("levene")
lev_result
print("Stats")
str(stat.test)
print("Significant?")
significance
print("Pairwise")
pwc


```


____________________________________________________________________
____________________________________________________________________

Scratch work : Don't look at this




One-way plot
```{r}

```




```{r}
pwc_plot <- pwc %>% add_xy_position(x = "Sample")


ggplot(data_set, aes(x = "Sample", y = "Number", color = "Treatment")) + 
  stat_pvalue_manual(pwc_plot, hide.ns = TRUE) +
  labs(
    #subtitle = get_test_label(stat.test, detailed = TRUE),
    caption = get_pwc_label(pwc_plot)
  )
```



WORKING BOXPLOT

```{r}
pwc_plot <- pwc %>% add_xy_position(x = "Sample")

box_plot <- ggboxplot(data_set, x = "Sample", y = "Number", color = "Time_points") +
  stat_pvalue_manual(pwc_plot, hide.ns = TRUE) +
  labs(
    #subtitle = get_test_label(stat.test, detailed = TRUE),
    caption = get_pwc_label(pwc_plot),
   outlier.shape = 8, position = position_dodge(1))  #geom_jitter(shape = 16, position = position_jitter(0.2))
box_plot



```



__________________________________________________________________

```{r}
  group_norm <- data_set%>%
    group_by(Sample) %>%
    shapiro_test(Number) 
  gncheck <- group_norm$p
  for (i in seq_along(gncheck)){
    if (gncheck[i] >0.05){shapirogn_result = TRUE
    }else {
      shapirogn_result = FALSE
      break
      }
   }
```

```{r}
### Post hoc test given significant ANOVA
pwc <- data_set %>% tukey_hsd(Number ~ Sample)

## Pairwise analysis, adding x and y info for adding to a box plot

pwc < data_set%>%
  group_by(groupings[1], groupings[2])%>%
  

pwc <- data_set%>%
  tukey_hsd( Number ~ Sample) %>%
  add_xy_position(x = "Sample")

#### Graph labeling
### For one way
pwc_lab <- pwc %>% add_xy_position(x= "Sample")
  #plot + stat_pvalue_manual(pwc_lab,hide.ns = TRUE) +
  # labs(
  # subtitle = get_test_label(stat.test, detailed = TRUE),
  # caption = get_pxc_label(pwc_lab))
  

```


Pairwise comparisons

```{r}
t.model<- lm (Number ~ Sample*Time_points, data = data_set)
data_set%>%
  emmeans_test(
    Number ~ Sample,  p.adjust.method = "bonferroni",
    model = t.model)

```

```{r}
 data_set%>%
  pairwise_t_test(
  Number ~ Sample, 
   p.adjust.method = "bonferroni"
)
```


Pairwise analysis, adding x and y info for adding to a box plot
```{r}
pwc <- data_set%>%
  tukey_hsd( Number ~ Sample) %>%
  add_xy_position(x = "Sample")
```

One way ANOVA:
dependent = Number
Treatment = Sample
```{r}
oneway_aov <- data_set %>% anova_test(Number ~ Sample)


```

Two way ANOVA"
dependent = Number
Treatment1 = Sample
Treatment2 = Time_points
Must first check a few asssumptions
```{r}
### check for outliers
data_set %>% 
  group_by(Sample, Time_points)%>%
 identify_outliers(Number)

### Check normalty
twoway_model<- lm(Number ~ Sample*Time_points,
                  data = data_set)
shapiro_test(residuals(twoway_model))

### Check normality of groups
data_set%>%
  group_by(Sample, Time_points)%>%
  shapiro_test(Number)

### Homogeneity of variance

data_set %>% levene_test(Number ~ Sample*Time_points)


twoway_anova <- data_set %>% anova_test(Number ~ Sample * Time_points)

## can do post-hoc tests

```
Post-hoct tests
A significant two-way interaction indicates that the impact that one factor (e.g., education_level) has on the outcome variable (e.g., job satisfaction score) depends on the level of the other factor (e.g., gender) (and vice versa). So, you can decompose a significant two-way interaction into:

Simple main effect: run one-way model of the first variable at each level of the second variable,
Simple pairwise comparisons: if the simple main effect is significant, run multiple pairwise comparisons to determine which groups are different.
For a non-significant two-way interaction, you need to determine whether you have any statistically significant main effects from the ANOVA output. A significant main effect can be followed up by pairwise comparisons between groups







Box_plot with p's
```{r}
aov <- oneway_aov
ggboxplot(
  data_set, x = "Time_points", y = "Number",
  color = "Sample", palette = "jco") +
  stat_pvalue_manual(pwc, hide.ns = TRUE)+ 
  labs (
  subtitle = get_test_label(aov, detailed = TRUE))

```

















```{r}
   melt_long_data <- melt(long_data, id.vars = c('Sample','Time_points'), measure.vars = 'Number')%>%
  print()
    
  ggplot(melt_long_data, aes(x = Time_points, y = value , color = Sample, group = Sample))+
           geom_line(linetype = "solid") + geom_point() + scale_color_viridis(discrete = TRUE, option = "D")
    
```








```{r}


tests <- long_data %>%
  group_by(Time_points)%>%
  t_test()
 
```

Pairwise comparisons

```{r}
 t.model<- lm (Number ~ Sample*Time_points, data = long_data) 
long_data%>%
  emmeans_test(
    Number ~ Sample,  p.adjust.method = "bonferroni",
    model = t.model)
 long_data%>%
  pairwise_t_test(
  Number ~ Sample, 
   p.adjust.method = "bonferroni"
)
```

ANOVA
```{r}
long_data.aov <- aov(formula = Number ~ Sample + Time_points, data = long_data )
summary(long_data.aov)
tukey_hsd(long_data.aov)


```

boxplot with T-test values
```{r}
pwc <- long_data%>%
  tukey_hsd( Number ~ Sample) %>%
  add_xy_position(x = "Sample")
aov <- long_data %>% anova_test(Number ~ Sample)
 ggboxplot(
  long_data, x = "Time_points", y = "Number",
  color = "Sample", palette = "jco") +
  stat_pvalue_manual(pwc, hide.ns = TRUE)+ 
  labs (
  subtitle = get_test_label(aov, detailed = TRUE))


```
One way anova
 https://www.datanovia.com/en/lessons/anova-in-r/#post-hoc-tests
 
identify_outliers()

get_summary_stats()

shaoriro test of normality
  model <- lm( y ~ x, data = )
  shapiro_test(residuals(model)) want p > 0.05 for noraml dist
 levene_test ( y~x) homogenety of variances
 
ANOVA:
 significant p's highlighted with *
 pairwise comparisons: tukey_hsd()


### visualize p-values:
see the reference link

look at effect of "Sample" on eveey time point. Two way anova

summary stats
```{r}
sum_stats <- long_data %>%
  group_by(Sample) %>%
  get_summary_stats(Number, type = "mean_sd") %>%
  print()
```


identify outlines
```{r}
long_data %>%
  group_by(Sample) %>%
  identify_outliers(Number)

```
build linear model
```{r}
data_lm<- lm(Number ~ Sample, 
             data = long_data)
long_data %>%
  group_by(Time_points)%>%
  anova_test(Number ~ Sample, error = data_lm)


```



Shapiro -Wilk test of normality

```{r}
shapiro_test(residuals(data_lm))
```

