---
title: "hw_05"
output: github_document
---

This homework is practice for splitting and grouping data frames


The code below imports libraries and data to be processed, and checks the structure of the data to be analyzed
```{r}
library(nycflights13)
library(tidyverse)

?nycflights13::flights

```

Gathers flights that were more than 2hrs later departing into a new data frame
```{r}
##gather flights that are at least 2hrs late
flights_2hr_late <- filter(flights, dep_delay >= 120)
print(flights_2hr_late)
```

Creates and new data frame with airline and flight number combined in a new column
```{r}
##flights with combined airline and flight number
flights_airline_num <- mutate(flights, airline_flight_num = paste0(carrier, flight))
print(flights_airline_num)
```

Created a new data frame displaying flights per carrier for each tailnumber. 
This is done by first grouping the flight data by tailnumber, then by carrier and creates a new data frame desplaying the group size of tailnumber by carrier
```{r}
## group flight data by tailnumber then by carrier
tail_num_group <- group_by(flights, tailnum, carrier)

## make tibble showing flights per carrier for each tailnumber aka number_of_flights
tail_num_carriers = summarize(tail_num_group, number_of_flights = n() )
print (tail_num_carriers)
```

